Considering scale in multilayer animal social networks

### 2020-06-15 paper/manuscript.Rmd
more tidy

Our examinations is motivated by behavioural, landscape, and spatial ecology to
build a conceptual and analytical framework for scale-dependent multilayer
networks, and we apply this framework to a case study of caribou (*Rangifer
tarandus*) in Newfoundland, Canada [@Peignier_2019]. We use this case study to
address two principal questions associated with scale-dependent multilayer
networks. First, to what degree does the social scale of relationships, spatial
scale of landscapes, and temporal scale of analysis influence our ability to
interpret complex social systems? Second, can variation in the social, spatial,
and temporal scale improve predictions associated with biological differences in
seasonal resource availability and social association? We conclude with
perspectives for further development and application of scale in multilayer
networks with respect to phenotypes and individual fitness,
movement ecology and collective movement, and habitat selection
and space use.
<!-- TODO: we answer the first question in the discussion, but not the second-->

@@

<!-- TODO: units of spatial scale? --> 
<!-- TODO: citations? -->

@@

## Caribou socioecology
<!--Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on predicted social (visual and auditory perception), spatial
(habitat type), and temporal (seasonal) scales.--> Caribou are gregarious
ungulates with fission-fusion dynamics [@Lesmerises_2018] that display temporal
and spatial variation in social networks [@Peignier_2019]. In winter, caribou
dig holes in the snow, termed craters, to access forage [@Bergerud_1974].
Cratering is presumed to be a costly behaviour and as a result, caribou tend to
occupy and re-use craters once they are established. <!-- and there is
considerably less access to forage than when the landscape is snow-free.-->
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. Access to forage is seasonally heterogeneous depending on snow
cover and, in winter, variable distribution of craters on the landscape.

@@

### Spatial scale
<!-- TODO: Increased spatial scale of land cover corresponds to xyz -->
To assess the influence of spatial scale, we aggregated the land cover raster using the R package `grainchanger` [@Graham_2019] across a series
of scales (100 - 1000 m by steps of 100 m). The land cover raster was aggregated
using a modal moving window method using a circular window corresponding to the
above scales [@Graham_2019a]. We assume 30 m and 100 m represents fine-scale
decision making for caribou during foraging, while re-sampling at 500 m , 750 m
and 1000 m represents the scale at which caribou tend to select and avoid
habitat (Bastille-Rousseau et al. 2017). For each spatial scale, we generated a
multilayer network consisting of the social association between 
`r DT[, uniqueN(get(idcol))]` individuals across three habitat layers (open, forest and
forage) for the entire study period. Social associations of individuals were
defined by spatiotemporal overlap within 5 minutes and 50 m. We calculated graph
strength and multidegree for each habitat layer and spatial scale, and edge
overlap across the entire series of networks to determine proportion of total
edges observed in each combination of habitat layer and spatial scale.

@@

# Results
<!-- TODO: Add intro/general to results?-->

Individuals in multilayer networks became more connected as the spatial distance
threshold increased. Mean graph strength differed across habitat layers with
similar and higher graph strength in open and forage layers, and lower strength
in forest layers (Figure \@ref(fig:socres)). Individual graph strength increased
sharply between 5 m and 100 m in all three habitat classes followed by a
decrease in rate after 100 m. These results suggest the optimal social scale at
which groups should be assigned is likely somewhere between ~20 - 100 m based on
the rate at which network metrics stabilized. For ungulates, groups are often
assigned based on 50 m threshold (Lingle 2003; Lesmeries et al. 2018; Peignier
et al. 2019), suggesting that given similar quantities of data, ~50 m is likely
an appropriate social scale for generating social networks in caribou and
similar species.

@@

<!-- Figure: Time window -->
(ref:temp) Varying temporal scale in multilayer networks. Multilayer networks were generated of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland within ordinal time windows (20 ordinal sample periods of 35-36 days) between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]`. A) Edge overlap across time windows showed two periods of low overlap (time windows 2-5 and 12-15), and three periods of high overlap (time windows 1, 6-11, 15-20). B) Ordinal network layers showing unweighted edges between individuals within time windows. Like edge overlap, individuals were more connected in time window 1, 6-11, and 15-20, compared to 2-5 and 12-15. C) Graph strength within time windows showing individuals as thin grey lines and mean across individuals as thick grey lines. Graph strength peaked in time window 9, with noticeable higher strength than the following year.

@@

Multilayer networks can be built with different types of interactions, a differ
In contrast to multilayer networks formed by combining layers defined
by different types of interactions, multilayer networks with GPS data are limited to 
proximity based measures. An alternative which has been explored is using 
hidden markov models behavioural states, within which individuals are associating. [@Jones_2020]
-->

<!-- temp -->
These changes in the outcome
depending on how the time window is defined highlight how critical it is to
properly define the time window appropriately to address the questions and
hypotheses of interest.

## Synth

Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations [@Hughey_2018]. Biologging and wildlife tracking technology has
advanced to the point where relocations can be programmed to occur as frequently
as each second [@B_rger_2016], a practice which has been applied to questions of
collective movement [e.g @Strandburg_Peshkin_2015]. Meanwhile, remote sensing
technology, including unmanned aerial vehicles and satellites, can be used to
track the fine-scale movements of groups over large distances and time scales
[@Torney_2018]. Using technology to infer behaviour has therefore become
standard operating procedure for many species in many systems. Our ability to
collect and process remotely sensed data and infer social processes based on
animal movement provides an opportunity to use multilayer social network
analysis to answer complex questions about animal socioecology.

>>>>

Our examinations is motivated by behavioural, landscape, and spatial ecology to
build a conceptual and analytical framework for scale-dependent multilayer
networks, and we apply this framework to a case study of caribou (*Rangifer
tarandus*) in Newfoundland, Canada [@Peignier_2019]. We use this case study to
address two principal questions associated with scale-dependent multilayer
networks. First, to what degree does the social scale of relationships, spatial
scale of landscapes, and temporal scale of analysis influence our ability to
interpret complex social systems? Second, can variation in the social, spatial,
and temporal scale improve predictions associated with biological differences in
seasonal resource availability and social association? We conclude with
perspectives for further development and application of scale in multilayer
networks with respect to phenotypes and individual fitness,
movement ecology and collective movement, and habitat selection
and space use.

@@

@@

## Caribou socioecology
Caribou are gregarious
ungulates with fission-fusion dynamics [@Lesmerises_2018] that display temporal
and spatial variation in social networks [@Peignier_2019]. In winter, caribou
dig holes in the snow, termed craters, to access forage [@Bergerud_1974].
Cratering is presumed to be a costly behaviour and as a result, caribou tend to
occupy and re-use craters once they are established. <!-- and there is
considerably less access to forage than when the landscape is snow-free.-->
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. Access to forage is seasonally heterogeneous depending on snow
cover and, in winter, variable distribution of craters on the landscape.

@@

### Spatial scale
To assess the influence of spatial scale, we aggregated the land cover raster using the R package `grainchanger` [@Graham_2019] across a series
of scales (100 - 1000 m by steps of 100 m). The land cover raster was aggregated
using a modal moving window method using a circular window corresponding to the
above scales [@Graham_2019a]. We assume 30 m and 100 m represents fine-scale
decision making for caribou during foraging, while re-sampling at 500 m , 750 m
and 1000 m represents the scale at which caribou tend to select and avoid
habitat (Bastille-Rousseau et al. 2017). For each spatial scale, we generated a
multilayer network consisting of the social association between 
`r DT[, uniqueN(get(idcol))]` individuals across three habitat layers (open, forest and
forage) for the entire study period. Social associations of individuals were
defined by spatiotemporal overlap within 5 minutes and 50 m. We calculated graph
strength and multidegree for each habitat layer and spatial scale, and edge
overlap across the entire series of networks to determine proportion of total
edges observed in each combination of habitat layer and spatial scale.

@@

# Results
Individuals in multilayer networks became more connected as the spatial distance
threshold increased. Mean graph strength differed across habitat layers with
similar and higher graph strength in open and forage layers, and lower strength
in forest layers (Figure \@ref(fig:socres)). Individual graph strength increased
sharply between 5 m and 100 m in all three habitat classes followed by a
decrease in rate after 100 m. These results suggest the optimal social scale at
which groups should be assigned is likely somewhere between ~20 - 100 m based on
the rate at which network metrics stabilized. For ungulates, groups are often
assigned based on 50 m threshold (Lingle 2003; Lesmeries et al. 2018; Peignier
et al. 2019), suggesting that given similar quantities of data, ~50 m is likely
an appropriate social scale for generating social networks in caribou and
similar species.

@@

<!-- Figure: Time window -->
(ref:temp) Varying temporal scale in multilayer networks. Multilayer networks were generated of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland within ordinal time windows (20 ordinal sample periods of 35-36 days) between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]`. A) Edge overlap across time windows showed two periods of low overlap (time windows 2-5 and 12-15), and three periods of high overlap (time windows 1, 6-11, 15-20). B) Ordinal network layers showing unweighted edges between individuals within time windows. Like edge overlap, individuals were more connected in time windows 1, 6-11, and 15-20, compared to 2-5 and 12-15. C) Graph strength within time windows showing individuals as thin grey lines and mean across individuals as thick grey lines. Graph strength peaked in time window 9, with noticeable higher strength than the following year.

@@

Multilayer networks can be built with different types of interactions, a differ
In contrast to multilayer networks formed by combining layers defined
by different types of interactions, multilayer networks with GPS data are limited to 
proximity based measures. An alternative which has been explored is using 
hidden markov models behavioural states, within which individuals are associating. [@Jones_2020]

t bon

-->

<!-- put this in -->
Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations [@Hughey_2018]. Biologging and wildlife tracking technology has
advanced to the point where relocations can be programmed to occur as frequently
as each second [@B_rger_2016], a practice which has been applied to questions of
collective movement [e.g @Strandburg_Peshkin_2015]. Meanwhile, remote sensing
technology, including unmanned aerial vehicles and satellites, can be used to
track the fine-scale movements of groups over large distances and time scales
[@Torney_2018]. Using technology to infer behaviour has therefore become
standard operating procedure for many species in many systems. Our ability to
collect and process remotely sensed data and infer social processes based on
animal movement provides an opportunity to use multilayer social network
analysis to answer complex questions about animal socioecology.
